---
title: "Comparing Metrics of Semantic Relatedness"
subtitle: "Analysis of Semantic Similarity Between Target Objects Using Different Representational Similarity Metrics"
output:
  pdf_document:
    df_print: kable 
    fig_caption: yes
    toc: yes
  html_document:
    df_print: paged
    toc: yes
geometry: margin=.75in
editor_options:
  tidy: yes
---



```{r}
# add packages
library(reshape2)
library(ggplot2)
```

## **Feature Norms**

Data Source: McRae, Cree, Seidenberg & McNorgan (2005)
```{r}
FN =read.csv(file = 'FeatureNorms.csv')
head(FN)
```

```{r}
# extract relevant cols
cFN = FN[,c(1,2,3,6)]
names(cFN) = c("Concept", "Feature", "F_Type", "F_Class")
head(cFN,10)
```


```{r}
# extract relevant features based on prior research
ccFN = cFN[cFN$F_Class!='taxonomic',]
ccFN = ccFN[ccFN$F_Class!='encyclopaedic',]
head(ccFN,10)
```

```{r}
# contruct vector of target object names
obs = c('book','bottle','cat','dog','shoes','spoon')

# extract data for each object
book = ccFN[ccFN$Concept=='book',]
bottle = ccFN[ccFN$Concept=='bottle',]
dog = ccFN[ccFN$Concept=='dog',]
cat = ccFN[ccFN$Concept=='cat',]
shoes = ccFN[ccFN$Concept=='shoes',]
spoon = ccFN[ccFN$Concept=='spoon',]

# construct vector of features for each obj
fBook = as.vector(book[,2])
fBottle = as.vector(bottle[,2])
fDog = as.vector(dog[,2])
fCat = as.vector(cat[,2])
fShoes = as.vector(shoes[,2])
fSpoon = as.vector(spoon[,2])

# get the set of unique features
featureSet = unique(sort(c(fBook, fBottle, fDog, fCat, fShoes, fSpoon)))
head(featureSet)
```

```{r}
# construct a FNs x Objects boolean df
OxFN = data.frame("Book" = featureSet%in%fBook,   "Bottle" = featureSet%in%fBottle, 
                  "Cat" = featureSet%in%fCat,     "Dog"=featureSet%in%fDog, 
                  "Shoes" = featureSet%in%fShoes, "Spoon"= featureSet%in%fSpoon, 
                  row.names = featureSet)
head(OxFN,10)
```

```{r}
# get correlation matrix for FNs of target objs
OxFN.cor = cor(OxFN)
as.data.frame(OxFN.cor)
```

```{r}
# format data for visualization
cors <- OxFN.cor
cor_data <- reshape2::melt(cors, varnames = paste0("objects", 1:2), value.name = "correlation")
```

```{r}
FN.plot <- ggplot(cor_data, aes(objects1, objects2, fill = correlation)) +
  geom_tile(color = "white")+ ggtitle("Correlation across Objects for Feature Norms")+
  scale_fill_gradient2(low = "yellow", high = "firebrick3", mid = "orange", 
                       midpoint = 0, limit = c(-1,1), space = "Lab", name="R-Values")+
  theme_minimal()+ 
  theme(axis.title.x = element_blank(),axis.title.y = element_blank(),
        axis.text.x = element_text(angle = 45, vjust = 1, size = 9, hjust = 1),
        plot.title = element_text(hjust = 0.5, size=11, face='bold'))+
  geom_text(aes(objects1, objects2, label = round(correlation,3)), color = "black", size = 3)+
coord_fixed()
FN.plot
```

## **WordNet Path Similarity**

Data Source: Princeton University "About WordNet." WordNet. Princeton University. 2010. 

```{r}
# Read in WordNet path similarities and format data
WN =read.csv(file = 'cWordNetSimilarities.csv',row.names = 1)
head(WN)
WN.cor = as.matrix(WN)
as.data.frame(WN.cor)
```

```{r}
# contruct vector of target object names
obs = c('baby','bottle','book','cat','dog','hand','shoes','spoon')
```

```{r}
# format data for visualization
cors <- WN.cor
cor_data <- reshape2::melt(cors, varnames = paste0("objects", 1:2), value.name = "correlation")
```

```{r}
WN.plot <- ggplot(cor_data, aes(objects1, objects2, fill = correlation))+
  geom_tile(color = "white")+ ggtitle("Similarity Between Objects in WordNet")+
  scale_fill_gradient2(low = "yellow", high = "firebrick3", mid = "firebrick2",
                       midpoint = 0.175, limit = c(0.05,.3), space = "Lab", name="Path Similarity",na.value = "firebrick3")+
  theme_minimal()+ 
  theme(axis.title.x = element_blank(),axis.title.y = element_blank(),
        axis.text.x = element_text(angle = 45, vjust = 1, size = 9, hjust = 1),
        plot.title = element_text(hjust = 0.5, size=11, face='bold'))+
  #geom_text(aes(objects1, objects2, label = round(correlation,3)), color = "black", size = 3)+ # uncomment to show path lengths
coord_fixed()
WN.plot
```

## **Co-occurance in CHILDES**
direc
Data Source: Dr. Thomas Hills and Eva Jimenez

```{r}
CC = read.csv(file = 'cooccuranceCHILDES.csv')
head(CC)
```

```{r}
# extract relevant cols
cCC = CC[,]
objects = c('X','baby', 'book','bottle','cat','dog','hand','shoe','spoon')
getCols = colnames(CC)%in%objects
cCC=cCC[,getCols]
head(cCC,10)

# extract relevant rows
getRows = cCC$X%in%objects
cCC=cCC[getRows,]

# reformat df
rownames(cCC) = cCC$X
cCC = cCC[,-1]
cCC
```

```{r}
# Reflect matrix along its diagonal to get a symmetric matrix
combinedCC = cCC

for (i in c(1:8))
  {
  for (j in c(1:8))
    {
    if (i!=j)
      {
      combinedCC[i,j]=cCC[i,j]+cCC[j,i]}
      }
  }

```

```{r}
# format for visualization
cCC.cor = as.matrix(combinedCC)
cors <- cCC.cor
cor_data <- reshape2::melt(cors, varnames = paste0("objects", 1:2), value.name = "correlation")
```

```{r}
# plot 
CC.plot <- ggplot(cor_data, aes(objects1, objects2, fill = log(correlation))) +
  geom_tile(color = "white")+ ggtitle("Co-occurance in CHILDES")+
  scale_fill_gradient2(low = "yellow", high = "firebrick3", mid = "firebrick2", midpoint = 6, 
                       limit = log(c(1,4668)), space = "Lab", na.value = "white", name="Co-occurance\nFrequency") +
  theme_minimal()+ 
  theme(axis.title.x = element_blank(),axis.title.y = element_blank(),
        axis.text.x = element_text(angle = 45, vjust = 1, size = 9, hjust = 1),
        plot.title = element_text(hjust = 0.5, size=11, face='bold'))+
  geom_text(aes(objects1, objects2, label = round(correlation,3)), color = "black", size = 3) +
coord_fixed()
CC.plot
```


```{r}
# normalize by frequency of the words
coocFreq = colSums(combinedCC)
ncCC=combinedCC/coocFreq
ncCC = t(ncCC)/coocFreq
ncCC = ncCC*100000
ncCC
```

```{r}
# format for visualization
ncCC.cor = as.matrix(ncCC)
cors <- ncCC.cor
cor_data <- reshape2::melt(cors, varnames = paste0("objects", 1:2), value.name = "correlation")
```

```{r}
CC.plot <- ggplot(cor_data, aes(objects1, objects2, fill = log(correlation))) +
  geom_tile(color = "white")+ ggtitle("Co-occurance in CHILDES")+
  scale_fill_gradient2(low = "yellow", high = "firebrick3", mid = "firebrick2", midpoint = 2, 
                       limit = c(-2,5), space = "Lab", na.value = "white", name="Co-occurance\nFrequency") +
  theme_minimal()+ 
  theme(axis.title.x = element_blank(),axis.title.y = element_blank(),
        axis.text.x = element_text(angle = 45, vjust = 1, size = 9, hjust = 1),
        plot.title = element_text(hjust = 0.5, size=11, face='bold'))+
  geom_text(aes(objects1, objects2, label = round(correlation,3)), color = "black", size = 3) +
coord_fixed()
CC.plot
```

## **CHILDES: Transformed into Similarity Space Using SVD**

```{r}
#Transforming CHILDES: Ananya 
CC = read.csv(file = 'cooccuranceCHILDES.csv')
rownames(CC) = CC[,1]
CC = CC[,-1]
CC
objects = c('baby','bottle','book','cat','dog','hand','shoe','spoon')

# create matrix
var_idx = which(colnames(CC) %in% objects) # get object indices
c_mat = t(CC[var_idx,]) # objects x all_words
dim(c_mat) # check dimensions

# without SVD
CHILDES_corr = cor(c_mat, method = "spearman")
colnames(CHILDES_corr)= colnames(CC)[var_idx]
rownames(CHILDES_corr)= colnames(CC)[var_idx]
#write.csv(CHILDES_corr, file="ChildesSimilarities_Corr.csv")

# perform svd
c_mat.svd = svd(c_mat) 
u = c_mat.svd$u
vt = c_mat.svd$v
sin_vals = c_mat.svd$d

# get k
explained_var = cumsum(sin_vals^2/sum(sin_vals^2))
plot(explained_var)
k = 4
# first 4 singular values capture the highest variance and therefore the most information

# reconstruct original matrix for k most significant singular values
# principle component analysis (basically)
vt = vt[1:k,]
u = u[,1:k]
sin_vals = diag(sin_vals[1:k])
t1 = u %*% sin_vals
t2 = t1 %*% vt
reconstructed_cmat = t2

# get correlation matrix
CHILDES_SVD_corr = cor(reconstructed_cmat, method = "spearman")
colnames(CHILDES_SVD_corr)= colnames(CC)[var_idx]
rownames(CHILDES_SVD_corr)= colnames(CC)[var_idx]
#write.csv(CHILDES_corr, file="ChildesSimilarities_SVD_4.csv")
CHILDES_SVD_corr
```

```{r}
# format data for visualization
cors <- CHILDES_SVD_corr
cor_data <- reshape2::melt(cors, varnames = paste0("objects", 1:2), value.name = "correlation")
```

```{r}
CC_SVD.plot <- ggplot(cor_data, aes(objects1, objects2, fill = correlation)) +
  geom_tile(color = "white")+ ggtitle("Similarity Between Objects in CHILDES, with SVD (k=4)")+
  scale_fill_gradient2(low = "yellow", high = "firebrick3", mid = "firebrick2", 
                       midpoint = 0.5, limit = c(-1,.99999999), space = "Lab", name="Correlations of \nCo-occurrence \nVectors", na.value = "firebrick3")+
  theme_minimal()+ 
  theme(axis.title.x = element_blank(),axis.title.y = element_blank(),
        axis.text.x = element_text(angle = 45, vjust = 1, size = 9, hjust = 1),
        plot.title = element_text(hjust = 0.5, size=11, face='bold'))+
  #geom_text(aes(objects1, objects2, label = round(correlation,3)), color = "black", size = 3)+
coord_fixed()
CC_SVD.plot
```


## **Neural Representation**

Data Source: Princeton Baby Lab Consortium Project fNRIS dataset 

```{r}
neural_rsm = read.csv(file = "mean_neural_RSM.csv", header = FALSE)
neural.cor = as.matrix(neural_rsm)
as.data.frame(neural.cor)
```

```{r}
# contruct vector of target object names
obs = c('baby','bottle','book','cat','dog','hand','shoes','spoon')
colnames(neural.cor) = obs
rownames(neural.cor) = obs
```

```{r}
# format data for visualization
cors <- neural.cor
cor_data <- reshape2::melt(cors, varnames = paste0("objects", 1:2), value.name = "correlation")
```

```{r}
neural.plot <- ggplot(cor_data, aes(objects1, objects2, fill = correlation))+
  geom_tile(color = "white")+ ggtitle("Similarity Between Objects Representations in the Brain")+
  scale_fill_gradient2(low = "yellow", high = "firebrick3", mid = "firebrick2",
                       midpoint = -.095, limit = c(-0.15,-0.04), space = "Lab", name="Correlations of \nChannel \nRepresentations", na.value = "firebrick3")+
  theme_minimal()+ 
  theme(axis.title.x = element_blank(),axis.title.y = element_blank(),
        axis.text.x = element_text(angle = 45, vjust = 1, size = 9, hjust = 1),
        plot.title = element_text(hjust = 0.5, size=11, face='bold'))+
  #geom_text(aes(objects1, objects2, label = round(correlation,3)), color = "black", size = 3)+
coord_fixed()
neural.plot
```



